---
title: "Workshop on Quarto"
author: "Samuel Pawel & Felix Hofmann"
bibliography: refs.bib
execute:
    echo: true
    warning: false
    message: false
# slighlty differnet YAML header specification than in R Markdown
# (e.g., "format" instead of "output", "html" instead of "html_document")
# format: odt
# format: pdf
format:
    html:
        toc: true
        number-sections: true
        colorlinks: true
# set "jupyter: python3" to use with python
---

<!-- R code chunks start as in R Markdown but different way to specify options  -->
```{r}
#| label: setup
#| include: false

# Load libraries
library(knitr)
library(ggplot2)
library(dplyr)
```

# Analysis of traffic accidents

In this section, we will quickly go through an example data set provided
by @ZHdata which contains data of traffic accidents in the city
of Zurich. In order to keep things simple, we have already pre-processed
the data and saved them into csv-files called
[accidents_2020.csv](https://raw.githubusercontent.com/crsuzh/dynamicReporting/presentation/workshop/accidents_2020.csv)
and
[accidents_2021.csv](https://raw.githubusercontent.com/crsuzh/dynamicReporting/presentation/workshop/accidents_2021.csv),
respectively.

## Importing the data

First of all, let's make sure, we can import the data sets into an `R`
session.

```{r}
#| label: data-import

# Set the name of the file we want to import
filename <- "../accidents_2021.csv"
# Read the data file
data <- read.csv(filename, header = TRUE, stringsAsFactors = FALSE)
```

Now that we have imported the data, we can have a look at the variables
within the data set. We can do this in two different ways.

First, we could list the names of the variables in the data set by
including an inline code chunk. We can achieve this by calling
`paste(names(data), collapse = ", ")`. Evaluating this generates
the following output: `r paste(names(data), collapse = ", ")`.

However, this does not tell us much yet. It would probably be more
helpful to actually see what this data looks like. 
Using the `kable` function in combination with the `head` function,
we can generate a table showing the first few observations of
the data table.

```{r}
#| label: tbl-display-data
#| tbl-cap: "The first six rows of our data set."

kable(head(data))
```

## Exploratory data analysis

Since the data is now in the form we want it to be, we can start
exploring it.

We could for example investigate how many accidents
happened per month or per weekday. A visual approach to answer
this question is shown in @fig-accidents-by-DayMonth.
Specifically, @fig-accidents-by-DayMonth-1 shows the number of
accidents by month whereas @fig-accidents-by-DayMonth-2 illustrates
the number of accidents by weekday.

```{r}
#| label: fig-accidents-by-DayMonth
#| fig-cap: "Number of accidents in Zurich by month and weekday."
#| fig-subcap:
#|      - "Number of accidents per month and year."
#|      - "Number of accidents per weekday and year."

data %>%
mutate(
    month = factor(month.name[month], levels = month.name),
    year = factor(year)
) %>%
group_by(month, year) %>%
summarise(accidents = n(), .groups = "drop") %>%
ggplot(aes(x = month, y = accidents, fill = year)) +
geom_col(position = position_dodge2()) +
theme_minimal() +
labs(x = "Month", y = "# Accidents", fill = "Year") +
theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 45, hjust = 1)
)


data %>%
mutate(
    weekday_name = factor(
        weekday_name,
        levels = c(
            "Monday", "Tuesday", "Wednesday",
            "Thursday", "Friday", "Saturday", "Sunday"
        )
    )
) %>%
group_by(weekday_name, year) %>%
summarise(accidents = n(), .groups = "drop") %>%
ggplot(aes(x = weekday_name, y = accidents, fill = as.factor(year))) +
geom_col(position = position_dodge2()) +
theme_minimal() +
labs(x = "Weekday", y = "# Accidents", fill = "Year") +
theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 45, hjust = 1)
)
```
It is generally good practice to include details of your session
in the document because it contains information about the computing environment.
This is important, because software changes over time and the analysis, thus, might
not run anymore with future versions. With this information, it is possible to
set up an environment with the exact same software versions that the analysis was
run in, which, in the end, increases the reproducibility of the results.

```{r}
#| label: sessionInfo

sessionInfo()
```

