---
title: "Workshop on R Markdown"
author: "Samuel Pawel & Felix Hofmann"
bibliography: refs.bib
# change here to something else (https://bookdown.org/yihui/rmarkdown/output-formats.html)
# some formats (e.g., pdf) require LaTeX installation!
output: 
  html_document:
    toc: true
    number_sections: true
    fig_caption: true
    self_contained: true
---

<!-- this is an R code chunk -->
<!-- Shortcut: insert with Ctrl + Alt + I (Mac: Cmd + Option + I) -->
```{r "setup", include = FALSE}
## packages
library(knitr)
library(ggplot2)
library(dplyr)

opts_chunk$set(
    echo = TRUE, # display R code
    warning = FALSE, # don't display warning messages
    message = FALSE, # don't display package loading messages
    fig.align = "center", # center figures
    fig.height = 4
)
```

# Markdown syntax

For a detailed overview see <https://bookdown.org/yihui/rmarkdown/markdown-syntax.html>

Here are some important parts:

`# Section title`

# Section title

`## Subsection title`

## Subsection title

`This is **bold** and this is *italic*.`

This is **bold** and this is *italic*.

`* List item 1`

`* List item 2`
 
* List item 1
* List item 2

`1) Numbered list item 1`

`2) Numbered list item 2`

1) Numbered list item 1
2) Numbered list item 2


# Analysis of traffic accidents

In this section, we will quickly go through an example data set provided
by @ZHdata which contains data of traffic accidents in the city of
Zurich. In order to keep things simple, we have already pre-processed
the data and saved them into csv-files called
[accidents_2020.csv](https://raw.githubusercontent.com/crsuzh/dynamicReporting/presentation/workshop/accidents_2020.csv)
and
[accidents_2021.csv](https://raw.githubusercontent.com/crsuzh/dynamicReporting/presentation/workshop/accidents_2021.csv)
, respectively.

## Importing the data

First of all, let's make sure, we can import the data sets into an `R`
session.

```{r "data-import"}
# Set the name of the file we want to import
filename <- "../accidents_2021.csv"
# Read the data file
data <- read.csv(filename, header = TRUE, stringsAsFactors = FALSE)
```

Now that we have imported the data, we can have a look at the variables
within the data set. We can do this in two different ways.

First, we could list the names of the variables in the data set by
including an inline code chunk. We can achieve this by calling
`paste(names(data), collapse = ", ")`. Evaluating this generates the
following output: `r paste(names(data), collapse = ", ")`.

However, this does not tell us much yet. It would probably be more
helpful to actually see what this data looks like. Using the `kable`
function in combination with the `head` function, we can generate a
table showing the first few observations of the data.

```{r "tbl-display-data"}
kable(head(data), caption = "The first six rows of our data set.")
```

## Exploratory data analysis

Since the data is now in the form we want it to be, we can start
exploring it.

We can for example investigate how many accidents happened per month or
per weekday. A visual approach to answer this question is shown in the 
following figures.

```{r "fig-accidents-by-DayMonth"}
data %>%
mutate(
    month = factor(month.name[month], levels = month.name),
    year = factor(year)
) %>%
group_by(month, year) %>%
summarise(accidents = n(), .groups = "drop") %>%
ggplot(aes(x = month, y = accidents, fill = year)) +
geom_col(position = position_dodge2()) +
theme_minimal() +
labs(x = "Month", y = "# Accidents", fill = "Year") +
theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 45, hjust = 1)
)


data %>%
mutate(
    weekday_name = factor(
        weekday_name,
        levels = c(
            "Monday", "Tuesday", "Wednesday",
            "Thursday", "Friday", "Saturday", "Sunday"
        )
    )
) %>%
group_by(weekday_name, year) %>%
summarise(accidents = n(), .groups = "drop") %>%
ggplot(aes(x = weekday_name, y = accidents, fill = as.factor(year))) +
geom_col(position = position_dodge2()) +
theme_minimal() +
labs(x = "Weekday", y = "# Accidents", fill = "Year") +
theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 45, hjust = 1)
)
```

It is generally good practice to include details of your session
in the document because it contains information about the computing environment.
This is important, because software changes over time and the analysis, thus, might
not run anymore with future versions. With this information, it is possible to
set up an environment with the exact same software versions that the analysis was
run in, which, in the end, increases the reproducibility of the results.

```{r "sessionInfo"}
sessionInfo()
```
