---
title: "Analysis of names from newborn children in Zurich"
author: "Felix Hofmann, Samuel Pawel"
date: "30 November 2022"
format: html
embed-resources: true
---

```{r}
#| echo: FALSE
## this is the R chunk to specify the options
knitr::opts_chunk$set(echo = TRUE, # display R code
                      warning = FALSE, # don't display warning messages
                      message = FALSE, # don't display package loading messages
                      fig.align = "center", # center figures
                      out.width = "100%" # figures fill entire width
                      )
```

In this document we will give an example of dynamic reporting using a data set containing the names of newborn children in the city of Zurich since the year 1993. The data were downloaded from <https://data.stadt-zuerich.ch/dataset/bev_vornamen_baby_od3700>.

## Import data and perform manipulations

We will now import the data and perform some manipulations, such as rename the variables into English and computing the most common name by year and sex.


```{r}
#| label: data-manipulations
## load data
dat <- read.csv(file = "../data/namesZurich.csv")

## data manipulations
library(dplyr)
datCleaned <- dat %>%
    rename(year = StichtagDatJahr,
           name = Vorname,
           sex = SexLang,
           births = AnzGebuWir) %>%
    group_by(year, sex) %>%
    top_n(n = 1, wt = births)
```

Last year, the most common male name was `r filter(datCleaned, year == 2021, sex == "männlich")$name` whereas the most common female name was `r filter(datCleaned, year == 2021, sex == "weiblich")$name`.

## Data visualization

We will now visualize the most common name by year and sex.

```{r}
#| label: plot
library(ggplot2)
ggplot(data = datCleaned, aes(x = year, y = births, color = sex)) +
    geom_line(aes(group = name), alpha = 0.3, show.legend = FALSE) +
    geom_label(aes(label = name), alpha = 0.8, show.legend = FALSE, size = 2) +
    scale_color_brewer(palette = "Dark2") +
    scale_x_continuous(breaks = seq(1995, 2020, 5)) + 
    labs(x = "Year", y = "Number of births",
         title = "Most common name in Zurich for newborns by year") +
    theme_minimal() +
    theme(panel.grid.minor = element_blank())
```

## Software versions and computational environment

It is good practice to save details on software versions and computational environment when creating dynamic reports.

```{r}
#| label: info
sessionInfo()
```