---
title: "Analysis of traffic accidents in Zürich"
author: "Samuel Pawel & Felix Hofmann"
format: html
embed-resources: true # create a single file with all depdencies inside
---

This document shows an analysis of traffic accidents in Zürich. The data data were downloaded
from the open data repository from the city of Zürich 
(<https://data.stadt-zuerich.ch/dataset/sid_dav_strassenverkehrsunfallorte>).

```{r echo = FALSE}
## import data into R
data <- read.csv("accidents_2020.csv") # <-- change here the data file

## compute the number of accidents in total
nrAccidents <- nrow(data)
```

The data set contains information on `r nrAccidents` traffic accidents. The following plot
shows the number of accidents per month.

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height = 4}
## create plot of month vs. number of traffic accidents
library(dplyr)
library(ggplot2)
data |>
    group_by(month_name, year) |>
    summarise(accidents = n()) |>
    mutate(monthfac = factor(month_name, levels = month.name)) |>
    ggplot(aes(x = monthfac, y = accidents, fill = factor(year))) +
    geom_col(position = position_dodge2(), width = 0.5) +
    labs(x = "Month", y = "Number of traffic accidents", fill = "Year") +
    theme_minimal() +
    theme(legend.position = "top",
          axis.text.x = element_text(angle = 45, vjust = 0.5),
          panel.grid.major.x = element_blank())
```

We see that in March and April 2020 there were much fewer traffic accidents compared to the 
rest of the year. A possible explanation is the national lockdown due to the Covid-19 pandemic.


```{r}
## print R version and package versions for reproducibility
sessionInfo()
```

